---
title: 'CAPSTONE PROJECT'
output:
  html_document: default
  word_document: default
  pdf_document: default
---
<center> <h1> Explanatory Data Analysis (EDA) </h1> </center>
<center>  <h3> [Zeynep Burcu Cevik] </h2> </center>
<center> <h3> [DHA & 501138112] </h2> </center>
---

#### 0. Importing dataset 
```{r}


fwb<-read.csv("/Users/zeynepcevik/Desktop/Ryerson/CIND820_Fall/Financial_wellbeing/NFWBS_PUF_2016_data.csv", header=TRUE, sep = ",", stringsAsFactors = FALSE)

```

#### 1. Check the datatypes of the attributes.
```{r}
str(fwb[1:90])
str(fwb[91:180])
str(fwb[181:217])
#Since not all the datatypes of the attribute are displayed, I divided the dataset into three parts.
#However we will not use all of them, because most of the variables are used for some scores such as financial well-being, financial knowledge., etc.
```

#### 3. Construct a subset from this dataset
```{r}
library(dplyr);
subfwb<-data.frame(fpl=fwb$fpl, FWBscore=fwb$FWBscore, FSscore=fwb$FSscore, LMscore=fwb$LMscore, KHscore=fwb$KHscore, EMPLOY=fwb$EMPLOY, Military_Status=fwb$Military_Status, agecat=fwb$agecat, generation=fwb$generation, PPEDUC=fwb$PPEDUC, PPETHM=fwb$PPETHM, PPGENDER=fwb$PPGENDER, PPHHSIZE=fwb$PPHHSIZE,PPINCIMP= fwb$PPINCIMP, PPMARIT=fwb$PPMARIT, PPMSACAT=fwb$PPMSACAT, PPREG4=fwb$PPREG4, PPREG9=fwb$PPREG9, PPT01=fwb$PPT01, PPT25=fwb$PPT25, PPT612=fwb$PPT612, PPT1317=fwb$PPT1317, PPT18OV=fwb$PPT18OV)
#Eliminate the participants who refused to respond. 
#From the data description, among these variables, only in employment and military status attributes, there are non-respondents.
subfwb<-filter(subfwb, subfwb$EMPLOY !=99 & subfwb$Military_Status !=-1)

```

#### 4. Check the datatypes of the attributes.
```{r}

str(subfwb)
```

#### 5. Descriptive statistics of the attributes.
```{r}
summary(subfwb)
any(is.na(subfwb)) #To test whether there are NA's in the dataset or not
```

#### 6. Some boxplots-nueric variables
```{r}
boxplot(subfwb$FWBscore, subfwb$FSscore , data=subfwb)
boxplot(subfwb$LMscore, subfwb$KHscore, data=subfwb)

```

#### 7. Boxplots
```{r}
boxplot(subfwb$FWBscore~subfwb$PPEDUC)
boxplot(subfwb$FWBscore~subfwb$PPINCIMP)
boxplot(subfwb$FWBscore~subfwb$PPMARIT)
boxplot(subfwb$FWBscore~subfwb$PPETHM)
boxplot(subfwb$FWBscore~subfwb$agecat)
boxplot(subfwb$FWBscore~subfwb$EMPLOY)
boxplot(subfwb$FWBscore~subfwb$PPHHSIZE)
boxplot(subfwb$FWBscore~subfwb$PPREG4)
boxplot(subfwb$FWBscore~subfwb$FSscore)
boxplot(subfwb$FWBscore~subfwb$LMscore)
boxplot(subfwb$FWBscore~subfwb$KHscore)
boxplot(subfwb$FWBscore~subfwb$PPGENDER)

```
#### 7. Boxplots
```{r}
boxplot(subfwb$FWBscore~subfwb$fpl)
boxplot(subfwb$FWBscore~subfwb$Military_Status)
boxplot(subfwb$FWBscore~subfwb$PPT01)
boxplot(subfwb$FWBscore~subfwb$PPT25)

```
#### 8. Correlation of non-categorical variables by using Pearson Correlation
```{r}

subfwb_noncat<-data.frame(subfwb$FWBscore,subfwb$FSscore,subfwb$KHscore)
cor(subfwb_noncat,use = "complete.obs")

```
####9. Correlation of categorical variables by using Spearman Correlation
```{r}

subfwb_cat<-data.frame(subfwb$fpl,subfwb$LMscore,subfwb$EMPLOY, subfwb$Military_Status, subfwb$agecat, subfwb$generation, subfwb$PPEDUC, subfwb$PPETHM, subfwb$PPGENDER, subfwb$PPHHSIZE, subfwb$PPINCIMP, subfwb$PPMARIT, subfwb$PPMSACAT, subfwb$PPREG4, subfwb$PPREG9, subfwb$PPT01, subfwb$PPT25, subfwb$PPT612, subfwb$PPT1317, subfwb$PPT18OV)
round(cor(subfwb_cat[1:3],use = "complete.obs", method="spearman"),3)
round(cor(subfwb,use="complete.obs", method="spearman"),4)



```

####10. Histograms for Quantitative Variables
```{r}

hist(subfwb$FWBscore,breaks=20)
hist(subfwb$FSscore)
hist(subfwb$KHscore)


```
#### 11. Boxplots
```{r}
boxplot(subfwb$FSscore~subfwb$PPEDUC)
boxplot(subfwb$KHscore~subfwb$PPEDUC)
boxplot(subfwb$PPINCIMP~subfwb$PPETHM)
boxplot(subfwb$FSscore~subfwb$PPREG4)


```



